---
title: "Palgen cytoeffect PLMM 3000"
author: "Lucy Quirant"
date: "4/23/2019"
output: 
  html_document:
    code_folding: show
    highlight: tango
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#Goal of document:
Apply the Poisson Log-Normal Mixed Model of the cytoeffect package to the Palgen HIV dataset (2019).

#cytoeffect - 3 parts: 
A) logistic linear mixed model (LLMM); 
B) poisson log-normal mixed model (PLMM) - 2000 cell subset;
C) poisson log-normal mixed model (PLMM) - 3000 cell subset;
This is part C.

# Prerequisites

###Parse input parameters.

```{r parse_input_parameters}
ncells = 3000
seed = 0xdada
ncores = parallel::detectCores()
zenodo_url = "https://zenodo.org/record/3248781#.XQo-raeB2L9/files/"
cytof_data = "se_palgen2019vaccine.Rdata" #created with CytoGLMM
prefit_PLMM = paste0("cytoeffect_plmm_ncells_",ncells,".Rdata")
prefit_PLMM
```

###Install packages.

```{r install_packages, warning=FALSE, message=FALSE, eval=FALSE}
pkgs_needed = c("devtools","tidyverse","magrittr","SummarizedExperiment",
                "ggthemes","cowplot","RColorBrewer","broom","hexbin",
                "intergraph","igraph","ggnetwork","ggcorrplot","MASS",
                "parallel","dplyr")
letsinstall = setdiff(pkgs_needed, installed.packages())
if (length(letsinstall) > 0) {
  BiocManager::install(letsinstall)
}
devtools::install_github("ChristofSeiler/cytoeffect")
```

###Load packages.

```{r load_packages}
library("cytoeffect")
library("tidyverse")
library("magrittr")
library("SummarizedExperiment")
library("ggthemes")
library("cowplot")
library("RColorBrewer")
library("broom")
library("intergraph")
library("igraph")
library("ggnetwork")
library("ggcorrplot")
library("MASS")
library("parallel")
library("dplyr")
theme_set(theme_few())
```

# Load Data

###Download preprocessed data from Zenodo.

```{r download_zenodo}
rdata_filenames = c(cytof_data, prefit_PLMM)
for(filename in rdata_filenames)
    download.file(url = paste0(zenodo_url, filename), 
                  destfile = filename,
                  mode = "wb")
```

###Load ``SummarizedExperiment`` object from CytoGLMM workflow.

```{r load_summarized_experiment}
load(cytof_data)
exprs = assay(se_palgen2019vaccine)
sample_info = rowData(se_palgen2019vaccine)
sample_info_names = names(sample_info)
df_samples = cbind(as.data.frame(exprs), as.data.frame(sample_info))
df_samples %<>% as.tibble
protein_names = colData(se_palgen2019vaccine) %>% 
  as.data.frame %>% 
  dplyr::filter(type != "unused") %>%
  .$protein_name
```

# Fit Model

###Tally cell count.

```{r cell_count}
df_samples %>% group_by(term,donor) %>% tally %>% arrange(n)
df_samples %>% group_by(term,donor) %>% tally %>% arrange(desc(n))
```

###Subsample cells to a maximum number of cells per donor.

```{r subsample_cells}
if(nrow(df_samples) > ncells) {
  print(paste("subsampled to",ncells,"per donor"))
  set.seed(seed)
  # subsample depending on max cell count
  df_count = df_samples %>% group_by(donor) %>% tally() %>%
    mutate(nnew = ifelse(n > ncells,ncells,n))
  # create table with a data frame in one column
  df_nested = df_samples %>% group_by(donor) %>% nest() %>%
    left_join(df_count,by = "donor")
  # subsample per donor
  df_samples = df_nested %>%
    mutate(samp = map2(data, nnew, sample_n)) %>%
    dplyr::select(donor, samp) %>%
    unnest()
} else {
  print("no subsampling done")
}
```

###Tally cell count.

```{r cell_count_subsample}
df_samples %>% group_by(term,donor) %>% tally %>% 
  arrange(n)
df_samples %>% group_by(term,donor) %>% tally %>% 
  arrange(desc(n))
```

## HMC Sampling

Sample from posterior distribution (initialization details in paper).

```{r set_reference}
df_samples$term %<>% factor(levels = c("Post-prime",
                                              "Post-boost"))
```


```{r poisson_sampling}
if(file.exists(prefit_PLMM)) {
  load(file = prefit_PLMM)
} else {
  obj = cytoeffect::poisson_lognormal(df_samples, protein_names, 
                                      condition = "term", group = "donor",
                                      iter = 325, warmup = 200, 
                                      num_chains = ncores)
  save(obj,file = prefit_PLMM)
}
```

## HMC Diagnostics

Postprocessing of posterior samples. Traceplot of posterior samples.

```{r post_sampling, fig.wide=TRUE}
rstan::traceplot(obj$fit_mcmc, inc_warmup = FALSE, pars = c("beta[1,1]", "beta[1,2]", "beta[2,1]", "beta[2,2]", "beta[3,1]", "beta[3,2]", "beta[4,1]", "beta[4,2]", "beta[5,1]", "beta[5,2]", "beta[6,1]", "beta[6,2]", "beta[7,1]", "beta[7,2]", "beta[8,1]", "beta[8,2]", "beta[9,1]", "beta[9,2]", "beta[10,1]", "beta[10,2]", "beta[11,1]", "beta[11,2]", "beta[12,1]", "beta[12,2]", "beta[13,1]", "beta[13,2]", "beta[14,1]", "beta[14,2]", "beta[15,1]", "beta[15,2]"))
rstan::traceplot(obj$fit_mcmc, inc_warmup = FALSE, pars = c("beta[16,1]", "beta[16,2]", "beta[17,1]", "beta[17,2]", "beta[18,1]", "beta[18,2]", "beta[19,1]", "beta[19,2]", "beta[20,1]", "beta[20,2]", "beta[21,1]", "beta[21,2]", "beta[22,1]", "beta[22,2]", "beta[23,1]", "beta[23,2]", "beta[24,1]", "beta[24,2]", "beta[25,1]", "beta[25,2]", "beta[26,1]", "beta[26,2]", "beta[27,1]", "beta[27,2]", "beta[28,1]", "beta[28,2]", "beta[29,1]", "beta[29,1]", "beta[30,1]", "beta[30,2]", "beta[31,1]", "beta[31,2]"))
```

Some more MCMC diagnostics. According to empirically findings, Rhat > 1.1 is usually indicative of problems in the fit.

```{r mcmc_diagnostics}
pars = c("beta",
         "sigma","sigma_term","sigma_donor",
         "Cor","Cor_term","Cor_donor", "b_donor") 
#b_donor seems to be newly added
tb = summary(obj$fit_mcmc, 
             pars = pars)$summary %>% 
  as.tibble(rownames = "pars", .before = 1) %>% 
  dplyr::select(pars, n_eff, Rhat)
tb %<>% na.omit() # Stan fills upper triangle with zeros
tb %>% arrange(n_eff)
tb %>% arrange(desc(Rhat))
tb %>% summarize(min = min(n_eff), max = max(n_eff))
tb %>% summarize(min = min(Rhat), max = max(Rhat))
```
Rhat very high for some correlation coefficients.

Further explore why some standard deviations are so large.

Traceplots of marker standard deviations
```{r}
pars_str = "sigma_donor"
rstan::traceplot(obj$fit_mcmc, inc_warmup = FALSE, pars = pars_str)
```

```{r}
pars_str = c(
  "beta[1,1]", "beta[1,2]", "b_donor[1,1]", "b_donor[2,1]",
  "sigma[1]", "sigma[2]", "sigma_donor[2]", "sigma_donor[10]", "Cor[1,2]", "Cor[21,3]",  "sigma[10]",
  "sigma_term[1]", "sigma_term[2]", "Cor_term[1,2]", "Cor_term[15,12]", "Cor_term[10,1]", "sigma_term[10]",
  "sigma_donor[1]",  "Cor_donor[1,2]",  "Cor[24,1]"
  )
rstan::traceplot(obj$fit_mcmc, inc_warmup = FALSE, pars = pars_str, ncol = 4)
```

```{r}
pars_str = c(
  "beta[10,1]", "b_donor[2,1]", "sigma[1]", "Cor[1,2]", "sigma[10]",
  "Cor_term[1,2]", "sigma_term[10]", "sigma_donor[13]", "Cor_donor[1,2]",
  "Cor_term[1,4]", "Cor_term[1,24]", "Cor_term[2,1]", "Cor_term[26,1]", "Cor[24,1]"
  )
rstan::traceplot(obj$fit_mcmc, inc_warmup = FALSE, pars = pars_str, ncol = 4)
```

# Results

###Plot posterior regression coefficients.

```{r plot_beta}
p1 = plot(obj, type = "beta") + 
  ggtitle(expression("Fixed Effects"~beta)) +
  theme(legend.position = "bottom") +
  guides(col = guide_legend(ncol = 1)) + 
  scale_color_few()
p1
plot(obj, type = "beta") + 
  facet_wrap(~condition, scales = "free_x") +
  theme(legend.position = "bottom") +
  guides(col = guide_legend(ncol = 1)) + 
  scale_color_few()
```

###Plot markers of interest from regression coefficients

```{r}
marker_selected = c("perforine", "HLADR", "GranzymeB", "CD66", "CD56", "CD45", "CD25", "CD2", "CD11a", "CD107a", "CCR5")
plot(obj, type = "beta", selection = marker_selected)

marker_selected_less = c("perforine", "HLADR", "GranzymeB", "CD66", "CD56", "CD45", "CD107a", "CCR5")
plot(obj, type = "beta", selection = marker_selected_less)

marker_selected_less = c("perforine", "CD66", "CD45")
plot(obj, type = "beta", selection = marker_selected_less)
ggsave(filename = "posterior_beta_significant_plmm_3000.pdf", width = 8, height = 6)
```

###Extract expected count difference for CD66

```{r fixed_effects_cd66}
post_beta = rstan::extract(obj$fit_mcmc, pars = "beta")[[1]]

prime_index = which(levels(pull(obj$df_samples, obj$condition)) 
                    == "Post-prime")
boost_index = which(levels(pull(obj$df_samples, obj$condition)) 
                    == "Post-boost")
cd66_index = which(obj$protein_names == "CD66")
prime_log_count = quantile(post_beta[,cd66_index,prime_index], 
                           probs = c(0.025, 0.5, 0.975))
prime_log_count
exp(prime_log_count)
boost_log_count = quantile(post_beta[,cd66_index,boost_index], 
                           probs = c(0.025, 0.5, 0.975))
boost_log_count
exp(boost_log_count)
diff_log_count = quantile(
  post_beta[,cd66_index,boost_index] - post_beta[,cd66_index,prime_index], 
  probs = c(0.025, 0.5, 0.975))
diff_log_count
exp(diff_log_count)
```

###Extract expected count difference for CD56

```{r fixed_effects_cd56}
post_beta = rstan::extract(obj$fit_mcmc, pars = "beta")[[1]]

prime_index = which(levels(pull(obj$df_samples, obj$condition)) 
                    == "Post-prime")
boost_index = which(levels(pull(obj$df_samples, obj$condition)) 
                    == "Post-boost")
cd56_index = which(obj$protein_names == "CD56")
prime_log_count = quantile(post_beta[,cd56_index,prime_index], 
                           probs = c(0.025, 0.5, 0.975))
prime_log_count
exp(prime_log_count)
boost_log_count = quantile(post_beta[,cd56_index,boost_index], 
                           probs = c(0.025, 0.5, 0.975))
boost_log_count
exp(boost_log_count)
diff_log_count = quantile(
  post_beta[,cd56_index,boost_index] - post_beta[,cd56_index,prime_index], 
  probs = c(0.025, 0.5, 0.975))
diff_log_count
exp(diff_log_count)
```

###Extract expected count difference for perforine

```{r fixed_effects_cd56}
post_beta = rstan::extract(obj$fit_mcmc, pars = "beta")[[1]]

prime_index = which(levels(pull(obj$df_samples, obj$condition)) 
                    == "Post-prime")
boost_index = which(levels(pull(obj$df_samples, obj$condition)) 
                    == "Post-boost")
perforine_index = which(obj$protein_names == "perforine")
prime_log_count = quantile(post_beta[,perforine_index,prime_index], 
                           probs = c(0.025, 0.5, 0.975))
prime_log_count
exp(prime_log_count)
boost_log_count = quantile(post_beta[,perforine_index,boost_index], 
                           probs = c(0.025, 0.5, 0.975))
boost_log_count
exp(boost_log_count)
diff_log_count = quantile(
  post_beta[,perforine_index,boost_index] - post_beta[,perforine_index,prime_index], 
  probs = c(0.025, 0.5, 0.975))
diff_log_count
exp(diff_log_count)
```

###Posterior multivariate pairs plot.

```{r posterior_pair_plot}
CD66_index = which(obj$protein_names == "CD66")
CD56_index = which(obj$protein_names == "CD56")
CD45_index = which(obj$protein_names == "CD45")

post_beta = rstan::extract(obj$fit_mcmc, pars = "beta")[[1]]
tb_log_count = bind_rows(
  tibble(
    term = levels(pull(obj$df_samples, obj$condition))[1],
    CD66 = post_beta[,CD66_index,1],
    CD56 = post_beta[,CD56_index,1],
    CD45 = post_beta[,CD45_index,1]
  ),
  tibble(
    term = levels(pull(obj$df_samples, obj$condition))[2],
    CD66 = post_beta[,CD66_index,2],
    CD56 = post_beta[,CD56_index,2],
    CD45 = post_beta[,CD45_index,2]
  )
)
plot_diag = function(marker) {
  ggplot(tb_log_count, aes_string(marker, fill = "term")) + 
    geom_histogram(bins = 40, position = "identity", alpha = 0.5) +
    scale_fill_few()
}
plot_off_diag = function(marker1, marker2) {
  ggplot(tb_log_count, aes_string(marker1, marker2, color = "term")) +
    geom_density2d() + 
    scale_color_few()
}
ppair = plot_grid(
  plot_diag("CD66") + theme(legend.position = "none"),
  NULL, 
  NULL,
  plot_off_diag("CD66","CD56") + theme(legend.position = "none"), 
  plot_diag("CD56") + theme(legend.position = "none"), 
  NULL,
  plot_off_diag("CD66","CD45") + theme(legend.position = "none"), 
  plot_off_diag("CD56","CD45") + theme(legend.position = "none"), 
  plot_diag("CD45") + theme(legend.position = "none"),
  ncol = 3
)
plot_grid(ppair,
          get_legend(plot_diag("CD66") + theme(legend.position = "bottom")),
          ncol = 1,
          rel_heights = c(1, .1))
ggsave(filename = "posterior_multivariate_plmm_3000.pdf", width = 8, height = 6)
```

###Plot posterior standard deviation.

```{r posterior_sigma}
p2 = plot(obj, type = "sigma") + 
  ggtitle("Marker Standard Deviation"~sigma) +
  theme(legend.position = "bottom") +
  guides(col = guide_legend(ncol = 1)) +
  scale_color_manual(values=c("#5DA5DA", "#FAA43A", "#F17CB0"))
p2
ggsave(filename = "posterior_sigma_3000.pdf")
```

###Plot posterior correlations.

```{r posterior_cor}
plist = plot(obj, type = "Cor")
plist
```

```{r}
plot.cytoeffect_poisson = function(obj, type = "beta",
                                   selection = obj$protein_names) {

  warmup = obj$fit_mcmc@stan_args[[1]]$warmup
  protein_names = obj$protein_names
  conditions = levels(pull(obj$df_samples_subset, obj$condition))
  
  if (type == "Cor") {

    var_names = c("Cor","Cor_term","Cor_donor")
    display_names = c(conditions,"donor")
    lapply(1:length(var_names), function(i) {
      cor = rstan::extract(obj$fit_mcmc, pars = var_names[i])[[1]]
      cor_median = apply(X = cor, MARGIN = c(2,3), FUN = median)
      colnames(cor_median) = rownames(cor_median) = protein_names
      ggcorrplot(cor_median, hc.order = TRUE, type = "lower",
                 outline.col = "lightgray",
                 colors = c("#6D9EC1", "white", "#E46726")) +
        ggtitle(paste0("Marker Correlations (",display_names[i],")")) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank())
    })

  } 

}
plot(obj, type = "Cor")


var_names = c("Cor","Cor_term","Cor_donor")
    display_names = c(conditions,"donor")
    lapply(1:length(var_names), function(i) {
      cor = rstan::extract(obj$fit_mcmc, pars = var_names[i])[[1]]
      cor_median = apply(X = cor, MARGIN = c(2,3), FUN = median)
      colnames(cor_median) = rownames(cor_median) = protein_names
      ggcorrplot(cor_median, hc.order = TRUE, type = "lower",
                 outline.col = "lightgray",
                 colors = c("#6D9EC1", "white", "#E46726")) +
        ggtitle(paste0("Marker Correlations (",display_names[i],")")) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank())
    })

```

```{r}
var_names = c("Cor","Cor_term","Cor_donor")
display_names = c(conditions,"donor")
titles = c("Post-prime", "Post-boost", "Donor")
lapply(1:length(var_names), function(i) {
      cor = rstan::extract(obj$fit_mcmc, pars = var_names[i])[[1]]
      cor_median = apply(X = cor, MARGIN = c(2,3), FUN = median)
      colnames(cor_median) = rownames(cor_median) = protein_names
      ggcorrplot(cor_median, hc.order = TRUE, type = "lower",
                 outline.col = "lightgray",
                 colors = c("#6D9EC1", "white", "#E46726")) +
        ggtitle(paste0("Marker Correlations (",titles[i],")")) +
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.text.x=element_text(size=10, angle=60),
              axis.text.y=element_text(size=10)) })

```

###Multivariate posterior MDS plots.

```{r multivariate_posterior}
cytoeffect::plot_mds(obj)
ggsave(filename = "posterior_mds_latent_variable_mu_scaled_3000.pdf", width = 10, height = 5)
cytoeffect::plot_mds(obj, asp = FALSE)
ggsave(filename = "posterior_mds_latent_variable_mu_unscaled_3000.pdf", width = 10, height = 5)
```

###Pairwise correlation change between conditions.

```{r correlation_uncertainty}
marker_pair = c("TNFa","IFNg")
Cor = rstan::extract(obj$fit_mcmc, pars = "Cor")[[1]]
Cor_term = rstan::extract(obj$fit_mcmc, pars = "Cor_term")[[1]]
Cor_diff = Cor_term - Cor
tb_cor = Cor_diff[,
                  which(obj$protein_names == marker_pair[1]),
                  which(obj$protein_names == marker_pair[2])] %>% as.tibble
tb_cor %<>% mutate(
  side = if_else(tb_cor$value > 0, 
                 true = paste0("positive (", 100*mean(tb_cor$value > 0), "%)"),
                 false = paste0("negative (", 100*mean(tb_cor$value <= 0), "%)"))
)
# keep colors consistent
if(mean(tb_cor$value > 0) == 1) {
  fill_colors = "#E46726"
} else {
  fill_colors = c("#6D9EC1","#E46726")
}
ggplot(tb_cor, aes(value, fill = side)) + 
  geom_histogram(bins = 50, alpha = 0.7) +
  xlab(paste0("Cor_term(", paste(marker_pair, collapse = ", "),")" )) +
  ggtitle("Posterior Distribution") + 
  scale_fill_manual(values = fill_colors)

#try different pairs
marker_pair = c("CD66","CD45")
tb_cor = Cor_diff[,
                  which(obj$protein_names == marker_pair[1]),
                  which(obj$protein_names == marker_pair[2])] %>% as.tibble
tb_cor %<>% mutate(
  side = if_else(tb_cor$value > 0, 
                 true = paste0("positive (", 100*mean(tb_cor$value > 0), "%)"),
                 false = paste0("negative (", 100*mean(tb_cor$value <= 0), "%)"))
)
ggplot(tb_cor, aes(value, fill = side)) + 
  geom_histogram(bins = 50, alpha = 0.7) +
  xlab(paste0("Cor_term(", paste(marker_pair, collapse = ", "),")" )) +
  ggtitle("Posterior Distribution") + 
  scale_fill_manual(values = fill_colors)
```

###Check if overall correlation structure changes between conditions.

```{r compare_covariance}
value = sapply(1:nrow(Cor_diff), function(i) {
  mask = which(upper.tri(Cor_diff[i,,]), arr.ind = T)
  cord = Cor_diff[i,,]
  mean(cord[lower.tri(cord)] > 0)
})
tb_cor = tibble(value = value)
tb_cor %<>% mutate(
  side = if_else(tb_cor$value > 0.5, 
                 true = paste0("> 1/2 (", 100*mean(tb_cor$value > 0.5), "%)"),
                 false = paste0("<= 1/2 (", 100*mean(tb_cor$value <= 0.5), "%)"))
)
p_global = ggplot(tb_cor, aes(value, fill = side)) + 
  geom_histogram(bins = 25, alpha = 0.7) +
  ggtitle(expression("Overall P(Corr"~Omega~"(3rd) > Corr"~Omega~"(1st))")) +
  scale_fill_manual(values = fill_colors) +
  theme(legend.position = "bottom") +
  xlab("probability")
p_global
```

###Plot differential correlations.

```{r plot_differential_cor}
cor_increase = apply(X = Cor_diff, MARGIN = c(2,3), FUN = function(x) mean(x > 0))
colnames(cor_increase) = rownames(cor_increase) = obj$protein_names
p_local = ggcorrplot(cor_increase, hc.order = TRUE, type = "lower",
           outline.col = "lightgray",
           colors = c("#6D9EC1", "white", "#E46726")) +
  ggtitle(expression("P(Corr"~Omega~"(Boost) > Corr"~Omega~"(Prime))")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  scale_fill_gradient2(limit = c(0, 1), midpoint = 0.5,
                       low = "#6D9EC1", mid =  "white", high = "#E46726",
                       name = "probability")
p_local
```

###Plot graph with edges at least 95% probability of larger correlation.

```{r plot_graph}
plot_correlation_graph = function(lambda) {
  graph = apply(Cor_diff, c(2,3), function(x) mean(x>0))
  diag(graph) = 0
  graph[upper.tri(graph)] = 0
  ind = which(graph > 0, arr.ind = T)
  tb_graph = tibble(
    from = obj$protein_names[ind[,1]],
    to = obj$protein_names[ind[,2]],
    prob = graph[ind]
  )
  tb_graph %<>% dplyr::filter(prob > lambda)
  tb_graph
  bayesFDR = sum(1-tb_graph$prob)/nrow(tb_graph)*100
  bayesFDR
  set.seed(0xdada)
  ig = graph_from_data_frame(tb_graph, directed = FALSE)
  ggplot(ggnetwork(ig, layout = "circle"), aes(x, y, xend = xend, yend = yend)) + 
    geom_edges(color = "black", size = 1) +
    geom_nodes(color = "black", size = 20) + 
    geom_nodetext(aes(label = vertex.names), 
                  color = "white", size = 3, fontface = "bold") +
    xlim(c(-0.1, 1.1)) +
    ylim(c(-0.1, 1.1)) + 
    ggtitle(paste0("Posterior Expected FDR: ", 
                   round(bayesFDR, digits = 1),"%")) +
    theme_blank() +
    theme(plot.title = element_text(hjust = 0.5))
}
plot_correlation_graph(lambda = 0.95)
plot_correlation_graph(lambda = 0.9)
plot_correlation_graph(lambda = 0.8)
plot_correlation_graph(lambda = 0.7)
```

###Combine plot for paper.

```{r compbine_plot}
pall = plot_grid(
  p1, p2, 
  plist[[1]] + ggtitle(expression("Marker Corr"~Omega~"(Post-prime)")),
  plist[[2]] + ggtitle(expression("Marker Corr"~Omega~"(Post-boost)")),
  p_global, p_local, 
  rel_heights = c(0.38,0.31,0.31),
  nrow = 3, labels = "AUTO"
)
ggsave(plot = pall, 
       filename = "posterior_summary_plmm_3000.pdf", 
       width = 8, height = 11)
```

# Goodness of Fit

Define a test statistics and compare observed value with posterior predictive distribution. 

###Predictive distribution marginalized over cell random effects.

```{r marginal_predicted_counts}
stan_pars = rstan::extract(obj$fit_mcmc, 
                           pars = c("beta",
                                    "sigma","sigma_term","sigma_donor",
                                    "Cor","Cor_term","Cor_donor"))
condition = "term"
term = obj$df_samples %>%
    pull(condition) %>%
    as.factor() %>%
    as.integer()
conditions_levels = levels(pull(obj$df_samples, 
                                obj$condition))
# kth posterior draw
sample_y_hat = function(k = 1) {
  set.seed(seed)
  lapply(1:2, function(cond) {
    n_cells_cond = table(term)[cond]
    beta = stan_pars$beta[k,,]
    mu = rep(0, length(protein_names))
    beta_rep = sapply(beta[,cond], rep, n_cells_cond)
    if(cond == 1) {
      sigma = stan_pars$sigma[k,]
      Cor = stan_pars$Cor[k,,]
    } else {
      sigma = stan_pars$sigma_term[k,]
      Cor = stan_pars$Cor_term[k,,]
    }
    Cov = diag(sigma) %*% Cor %*% diag(sigma)
    b = mvrnorm(n = n_cells_cond, mu, Cov)
    sigma_donor = stan_pars$sigma_donor[k,]
    Cor_donor = stan_pars$Cor_donor[k,,]
    Cov_donor = diag(sigma_donor) %*% Cor_donor %*% diag(sigma_donor)
    b_donor = mvrnorm(n = n_cells_cond, mu, Cov_donor)
    count = exp(beta_rep + b + b_donor)
    count = matrix(rpois(length(count), count),
                   nrow = nrow(count), 
                   ncol = ncol(count))
    count %<>% as.tibble
    names(count) = protein_names
    count %<>% add_column(term  = conditions_levels[cond])
    count
  }) %>% bind_rows()
}
Y_hat = sample_y_hat(k = 1)
Y_hat %>% 
  group_by(term) %>% 
  summarize_at(protein_names, median)
```
The only one that is not O or NA is CD66 with post-boost being 1 and post-prime equal to 0.

Check if we can model CD66, perforine, and CD45 bright cells.

```{r goodness_of_fit_a}
gof = function(df, test_stat) {
  tfm = function(x) asinh(x/5)
  df_tfm = df %>% mutate_at(protein_names, tfm)
  df_median = df_tfm %>% summarize_at(protein_names, median)
  tibble(
    term = c(
      "Post-prime", 
      "Post-boost"
      ),
    statistic = c(
      test_stat(df_tfm %>% filter(term == "Post-prime"), df_median), 
      test_stat(df_tfm %>% filter(term == "Post-boost"), df_median)
      )
  )
}
test_stat_a = function(df_tfm, df_median) {
  mean(df_tfm$CD66 > df_median$CD66 & 
       df_tfm$perforine > df_median$perforine & 
       df_tfm$CD45 > df_median$CD45) * 100
}
gof_obsv_a = gof(df_samples, test_stat_a)
gof_pred_a = mclapply(1:dim(stan_pars$beta)[1], function(k) gof(sample_y_hat(k), test_stat_a),
                      mc.cores = ncores) %>% bind_rows() 
gof_obsv_a %<>% mutate(subset = "Cell Subset A")
gof_pred_a %<>% mutate(subset = "Cell Subset A")
```

Check if we can model CD66 bright, and perforine and CD45 dim cells.

```{r goodness_of_fit_b}
test_stat_b = function(df_tfm, df_median) {
 mean(df_tfm$CD66 > df_median$CD66 & 
       df_tfm$perforine < df_median$perforine & 
       df_tfm$CD45 < df_median$CD45) * 100
}
gof_obsv_b = gof(df_samples, test_stat_b)
gof_pred_b = mclapply(1:dim(stan_pars$beta)[1], function(k) gof(sample_y_hat(k), test_stat_b),
                      mc.cores = ncores) %>% bind_rows()
gof_obsv_b %<>% mutate(subset = "Cell Subset B")
gof_pred_b %<>% mutate(subset = "Cell Subset B")
```

Check if we can model zero CD56 and CD11a bright cells.

```{r goodness_of_fit_c}
test_stat_c = function(df_tfm, df_median) {
  mean(df_tfm$CD56  == 0 & 
       df_tfm$CD11a > df_median$CD11a) * 100
}
gof_obsv_c = gof(df_samples, test_stat_c)
gof_pred_c = mclapply(1:dim(stan_pars$beta)[1], function(k) gof(sample_y_hat(k), test_stat_c),
                      mc.cores = ncores) %>% bind_rows()
gof_obsv_c %<>% mutate(subset = "Cell Subset C")
gof_pred_c %<>% mutate(subset = "Cell Subset C")
```

Check if we can model nonzero CD56 and CD11a bright cells.

```{r goodness_of_fit_d}
test_stat_d = function(df_tfm, df_median) {
 mean(df_tfm$CD56 > 0 & 
       df_tfm$CD11a > df_median$CD11a) * 100
}
gof_obsv_d = gof(df_samples, test_stat_d)
gof_pred_d = mclapply(1:dim(stan_pars$beta)[1], function(k) gof(sample_y_hat(k), test_stat_d),
                      mc.cores = ncores) %>% bind_rows()
gof_obsv_d %<>% mutate(subset = "Cell Subset D")
gof_pred_d %<>% mutate(subset = "Cell Subset D")
```

Combined plot.

```{r goodness_of_fit_combined}
# combine observed statistic
gof_obsv_all = bind_rows(gof_obsv_a, gof_obsv_b, gof_obsv_c, gof_obsv_d)
# combined predicted statistic
gof_pred_all = bind_rows(gof_pred_a, gof_pred_b, gof_pred_c, gof_pred_d)
# plot everything
ggplot(gof_pred_all, aes(statistic, fill = term)) + 
  geom_histogram(bins = 40, position = "identity", alpha = 0.5) +
  geom_vline(data = gof_obsv_all, linetype = "dashed", size = 1, 
             aes(xintercept = statistic, color = term)) +
  scale_fill_few() + 
  scale_color_few() + 
  xlab("test statistic (percentage)") + 
  facet_wrap(~subset, scales = "free") + 
  theme(legend.position="bottom")
ggsave(filename = "goodness_of_fit.pdf", 
       width = 8, height = 5)
```

# Session Info {.unnumbered}

```{r session_info}
sessionInfo()
```