---
title: "Plagen cytoeffect LLMM"
author: "Lucy Quirant"
date: "4/23/2019"
output: 
  html_document:
    code_folding: show
    highlight: tango
    toc: true
    toc_depth: 3
    number_sections: true
---

#Goal of document:
Apply the Logistic Linear Mixed Model of the cytoeffect package to the Palgen HIV dataset (2019).

#cytoeffect - 2 parts: 
A) logistic linear mixed model (LLMM); 
B) poisson log-normal mixed model (PLMM); 
This is part A.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Prerequisites

###Parse input parameters.

```{r parse_input_parameters}
ncells = 10000 #later can set to Inf (infinite)
seed = 0xdada
ncores = parallel::detectCores()
cytof_data = "se_palgen2019vaccine.Rdata"
prefit_LLMM = paste0("cytoeffect_llmm_ncells_",ncells,".Rdata")
prefit_LLMM
prefit_LLMM_reduced = paste0("cytoeffect_llmm_ncells_",ncells,"_reduced.Rdata")
prefit_LLMM_reduced
prefit_LLMM_reduced_CD45 = paste0("cytoeffect_llmm_ncells_",ncells,"_reduced_CD45.Rdata")
prefit_LLMM_reduced_CD45
```

###Install packages.

```{r install_packages, warning=FALSE, message=FALSE, eval=FALSE}
pkgs_needed = c("devtools","tidyverse","magrittr","SummarizedExperiment",
                "ggthemes","cowplot")
letsinstall = setdiff(pkgs_needed, installed.packages())
if (length(letsinstall) > 0) {
  source("http://bioconductor.org/biocLite.R")
  biocLite(letsinstall)
}
devtools::install_github("ChristofSeiler/cytoeffect")
```

###Load packages.

```{r load_packages}
library("cytoeffect")
library("tidyverse")
library("magrittr")
library("SummarizedExperiment")
library("ggthemes")
library("cowplot")
theme_set(theme_few())
```

# Load Data

###Load ``SummarizedExperiment`` object from CytoGLMM workflow.

```{r load_summarized_experiment}
rdata_filenames = c(cytof_data, prefit_LLMM, prefit_LLMM_reduced, prefit_LLMM_reduced_CD45)
load(cytof_data)
exprs = assay(se_palgen2019vaccine)
sample_info = rowData(se_palgen2019vaccine)
sample_info_names = names(sample_info)
df_samples = cbind(as.data.frame(exprs), as.data.frame(sample_info))
df_samples %<>% as.tibble
protein_names = colData(se_palgen2019vaccine) %>% 
  as.data.frame %>% 
  dplyr::filter(type != "unused") %>%
  .$protein_name
```

Transform counts.

```{r transform_counts}
df_samples %<>% mutate_at(colData(se_palgen2019vaccine)$protein_name, 
                          function(x) asinh(x/5))
```

# Fit Model

###Tally cell count.

```{r cell_count}
df_samples %>% 
  group_by(term,donor) %>% 
  tally %>% 
  arrange(n)
df_samples %>% 
  group_by(term,donor) %>% 
  tally %>% 
  arrange(desc(n))
```

###Subsample cells to a maximum number of cells per donor.

```{r subsample_cells}
if(nrow(df_samples) > ncells) {
  print(paste("subsampled to",ncells,"per donor"))
  set.seed(seed)
  # subsample depending on max cell count
  df_count = df_samples %>% group_by(donor) %>% tally() %>%
    mutate(nnew = ifelse(n > ncells,ncells,n))
  # create table with a data frame in one column
  df_nested = df_samples %>% group_by(donor) %>% nest() %>%
    left_join(df_count,by = "donor")
  # subsample per donor
  df_samples = df_nested %>%
    mutate(samp = map2(data, nnew, sample_n)) %>%
    dplyr::select(donor, samp) %>%
    unnest()
} else {
  print("no subsampling done")
}
```

###Tally cell count.

```{r cell_count_subsample}
df_samples %>% group_by(term,donor) %>% tally %>% 
  arrange(n)
df_samples %>% group_by(term,donor) %>% tally %>% 
  arrange(desc(n))
```

## HMC Sampling

Sample from posterior distribution.

```{r set_reference}
df_samples$term %<>% factor(levels = c("Post-prime",
                                              "Post-boost"))
```

```{r logistic_sampling}
if(file.exists(prefit_LLMM)) {
  load(file = prefit_LLMM)
} else {
  obj = cytoeffect::glmm(df_samples, protein_names, 
                       condition = "term", group = "donor",
                       iter = 325, warmup = 200, 
                       num_chains = ncores)
  save(obj,file = prefit_LLMM)
}
```

## HMC Diagnostics

Traceplot of posterior samples.

```{r post_sampling, fig.wide=TRUE}
rstan::traceplot(obj$fit_mcmc, inc_warmup = TRUE)
rstan::traceplot(obj$fit_mcmc, inc_warmup = FALSE)
```

Some more MCMC diagnostics. According to empirically findings, Rhat > 1.1 is usually indicative of problems in the fit.

```{r mcmc_diagnostics}
pars = c("beta","sigma_donor","L_donor","z_donor")
tb = summary(obj$fit_mcmc, 
             pars = pars)$summary %>% 
  as.tibble(rownames = "pars", .before = 1) %>% 
  select(pars, n_eff, Rhat)
tb %<>% na.omit() # Stan fills upper triangle with zeros
tb %>% arrange(n_eff)
tb %>% arrange(desc(Rhat))
tb %>% summarize(min = min(n_eff), max = max(n_eff))
tb %>% summarize(min = min(Rhat), max = max(Rhat))
```

# Results

###Plot fixed effects.

```{r fixed_effects, fig.small=TRUE}
p_full = plot(obj, type = "beta") + 
  ggtitle(expression("Fixed Effects"~beta)) +
  xlab("log-odds of Post-boost/ Post-prime")
p_full
```

###Extract log-odds for CD66 

```{r fixed_effects_CD66}
post_beta = rstan::extract(obj$fit_mcmc, pars = "beta")[[1]]
post_beta %<>% as.tibble
names(post_beta) = c("intercept",obj$protein_names)
quantile(post_beta$CD66, probs = c(0.025, 0.975))
quantile(exp(post_beta$CD66), probs = c(0.025, 0.975))
```

###Plot random effects.

```{r random_effects}
plot(obj, type = "sigma_donor") + 
  ggtitle("Marker Standard Deviation"~sigma)
ggsave("sigma_glmm_10000_v2.pdf")
```

###Plot posterior correlations.

```{r posterior_correlations, fig.small=TRUE}
plot(obj, type = "Cor_donor") + 
  ggtitle(expression("Marker Correlations"~Omega~"(donor)"))
ggsave("posterior_summary_cor_glmm_10000.pdf", width = 4, height = 4)
```

# Refit Without CD66

###Refit model to test potentional collider confounding.

```{r refit}
protein_names = obj$protein_names[-which(obj$protein_names == "CD66")]
if(file.exists(prefit_LLMM_reduced)) {
  load(file = prefit_LLMM_reduced)
} else {
  obj = cytoeffect::glmm(df_samples, protein_names, 
                         condition = "term", group = "donor",
                         iter = 325, warmup = 200, 
                         num_chains = ncores)
  save(obj,file = prefit_LLMM_reduced)
}
```

```{r plot_refit}
p_reduced = plot(obj, type = "beta") + 
  ggtitle(expression("Reduced Fixed Effects"~beta)) +
  xlab("log-odds of Post-boost/ Post-prime")
plot_grid(p_full, p_reduced, labels = "AUTO")
ggsave(filename = "beta_glmm_10000_v2.pdf")
```

# Session Info {.unnumbered}

```{r session_info}
sessionInfo()
```
